---
title: "Lab 05: Candy"
subtitle: "Recipes + Model comparison"
execute:
  freeze: auto
  echo: true
  eval: false
  warning: false
  message: false
---

::: callout-important
Due:

-   Monday, October 24 , 11:59pm (Thursday labs)
-   Tuesday, October 25, 11:59pm (Friday labs)
:::

## Introduction

In today's lab you will analyze data on candy obtained from an online experiment conducted at [FiveThirtyEight](https://fivethirtyeight.com/).

### Learning goals

By the end of the lab you will be able to

-   describe the components of a recipe
-   fit a model using recipes
-   compare models
-   continue developing a collaborative workflow with your teammates

## Getting started

-   A repository has already been created for you and your teammates. Everyone in your team has access to the same repo.
-   Go to the [sta210-fa22](https://www.github.com/sta210-fa22) organization on GitHub. Click on the repo with the prefix **lab-05**. It contains the starter documents you need to complete the lab.
-   Each person on the team should clone the repository and open a new project in RStudio. Throughout the lab, each person should get a chance to make commits and push to the repo.
-   **Do not make any changes to the `.qmd` file until the instructions tell you do to so.**

## Workflow: Using Git and GitHub as a team

::: callout-important
There are no Team Member markers in this lab; however, you should use a similar workflow as in Lab 04. Only one person should type in the group's Qmd file at a time. Once that person has finished typing the group's responses, they should *render, commit, and push* the changes to GitHub. All other teammates can *pull* to see the updates in RStudio.

Every teammate must have at least one commit in the lab. Everyone is expected to contribute to discussion even when they are not typing.
:::

## Packages

The following packages are used in the lab.

```{r}
#| label: load-packages
#| message: false
#| eval: true

library(tidyverse)
library(tidymodels)
library(fivethirtyeight)
library(knitr)

# add other packages as needed
```

## Data: Candy

The data from this lab comes from the the article FiveThirtyEight [*The Ultimate Halloween Candy Power Ranking*](https://fivethirtyeight.com/features/the-ultimate-halloween-candy-power-ranking) by Walt Hickey. To collect data, Hickey and collaborators at FiveThirtyEight set up an experiment people could vote on a series of randomly generated candy matchups (e.g. Reeses vs. Skittles). Click [here](http://walthickey.com/2017/10/18/whats-the-best-halloween-candy/) to check out some of the match ups.

The data set contains the characteristics and win percentage from 85 candies in the experiment. The variables are

| Variable           | Description                                                               |
|---------------------------|---------------------------------------------|
| `chocolate`        | Does it contain chocolate?                                                |
| `fruity`           | Is it fruit flavored?                                                     |
| `caramel`          | Is there caramel in the candy?                                            |
| `peanutalmondy`    | Does it contain peanuts, peanut butter or almonds?                        |
| `nougat`           | Does it contain nougat?                                                   |
| `crispedricewafer` | Does it contain crisped rice, wafers, or a cookie component?              |
| `hard`             | Is it a hard candy?                                                       |
| `bar`              | Is it a candy bar?                                                        |
| `pluribus`         | Is it one of many candies in a bag or box?                                |
| `sugarpercent`     | The percentile of sugar it falls under within the data set. Values 0 - 1. |
| `pricepercent`     | The unit price percentile compared to the rest of the set. Values 0 - 1.  |
| `winpercent`       | The overall win percentage according to 269,000 matchups. Values 0 - 100. |

Use the code below to get a glimpse of the `candy_rankings` data frame in the **fivethirtyeight** R package.

```{r eval = F}
glimpse(candy_rankings)
```

## Exercises

The goal of this analysis is to multiple linear regression to understand the factors that make a good candy.

### Exercise 1

Split the data into training (80%) and test (20%) sets. Name the data sets `candy_train` and `candy_test`, respectively. Use `set.seed(1)`.

::: callout-tip
Use argument `prop =` in the `initial_split` function to specify the proportion in the training data.
:::

### Exercise 2

Below is a recipe for a model using the characteristics of candy to understand variability in the win percentage. Each line of the recipe code is labeled Line 1 - Line 8. Describe what each line of code does. The explanation should be written comprehensively and specifically enough that someone could replicate the data manipulation steps.

For example, if a line of code is `step_center(X)`, a comprehensive and specific explanation is like the following: "This line of code mean centers the variable $X$ by subtracting $\bar{X}$ from each value of $X$ in the training data."

::: callout-tip
Use the [Recipes Function Reference](https://recipes.tidymodels.org/reference/index.html) page as a resource to learn more about the recipe functions.
:::

```{r}
#| eval: false

#Line 1
candy_rec <- recipe(winpercent ~ ., data = candy_train) |> 
#Line 2
  update_role(competitorname, new_role = "ID") |> 
# Line 3
  step_cut(sugarpercent, breaks = c(0, 0.25, 0.5, 0.75,1)) |> 
#Line 4
  step_mutate(pricepercent = pricepercent * 100) |> 
#Line 5
  step_dummy(all_nominal_predictors()) |> 
#Line 6
  step_interact(terms =~ pricepercent:chocolate) |> 
#Line 7
  step_rm(fruity, caramel, peanutyalmondy, nougat, hard, bar, pluribus, cripsedricewafer) |> 
#Line 8
  step_zv() 
```

### Exercise 3

How many terms (not including the intercept) will be in the model produced by the recipe in Exercise 2?

### Exercise 4

Specify the model, build the model workflow using the recipe in Exercise 2, and fit the model to the training data. Neatly display the model using 3 digits.

```{r}
candy_spec <- linear_reg() |>
  set_engine("lm")

candy_workflow <- workflow() |>
  add_model(candy_spec) |>
  add_recipe(candy_rec) 

candy_fit <- candy_workflow |>
  fit(data = candy_train)

tidy(candy_fit) |>
  kable(digits = 3)
```

### Exercise 5

Interpret the following in the context of the data:

-   Intercept

-   Coefficient of `sugarpercent_X.0.75.1.`

-   Coefficient of `pricepercent_x_chocolateTRUE`

### Exercise 6

Let's consider another model. Use the `recipe()` workflow to fit a new model that meets the following criteria:

-   Includes variables `chocolate`, `pricepercent`, `cripsedricewafer`, `pluribus, sugarpercent`

-   Rescales `pricepercent` so it ranges 0 to 100 (instead of 0 to 1)

-   Makes `sugarpercent` a factor where the levels equal the four quartiles - 0 - 25, 25 - 50, 50 - 75, 75 - 100

-   Includes interaction between `pricepercent` and `pluribus`

Neatly display the model using 3 digits.

### Exercise 7

Calculate $R^2$, Adjusted $R^2$, AIC, and BIC for the models fit in Exercise 3 and the Exercise 6.

### Exercise 8

Consider the model from Exercise 3 "Model 1" and the model fit in Exercise 6 "Model 2". State the model you would select based on the criteria below. Briefly explain your response.

-   $R^2$

-   Adjusted $R^2$

-   AIC

-   BIC

### Exercise 9

Use RMSE to evaluate the predictive performance of each model on the testing data. Which model would you choose based on RMSE?

### Exercise 10

-   Which model do you choose after taking into account the model comparison results in Exercises 9 and 10? Briefly explain your response.

-   Use the model you selected to describe what generally makes a good candy, i.e., one with a high win percentage.

## Submission

::: callout-warning
Before you wrap up the assignment, make sure all documents are updated on your GitHub repo. We will be checking these to make sure you have been practicing how to commit and push changes.

Remember -- you must turn in a PDF file to the Gradescope page before the submission deadline for full credit.
:::

One team member submit the assignment:

-   Go to [http://www.gradescope.com](http://www.gradescope.com/) and log in using your NetID credentials.
-   Click on your *STA 210* course.
-   Click on the assignment, and you'll be prompted to submit it.
-   Select all team members' names, so they receive credit on the assignment. [Click here](https://help.gradescope.com/article/m5qz2xsnjy-student-add-group-members) for video on adding team members to assignment on Gradescope.
-   Mark the pages associated with each exercise. All of the pages of your lab should be associated with at least one question (i.e., should be "checked").
-   Select the first page of your PDF submission to be associated with the *"Workflow & formatting"* section.

## Grading

Total points available: 50 points.

| Component             | Points |
|-----------------------|--------|
| Ex 1 - 10             | 45     |
| Workflow & formatting | 5[^1]  |

[^1]: The "Workflow & formatting" grade is to assess the reproducible workflow. This includes having at least one meaningful commit from each team member and updating the team name and date in the YAML.
